My greatest thanks are to my special advisor Prof. Kaleem Siddiqi, who helped me during these years with the challenges I have faced. He made me feel supported all the time and he helped me in all areas of my graduate life gently. I strongly believe that working under his supervision is one of the very rare chances that a graduate student can be given and for this I am grateful.

I am deeply thankful to all the shape analysis group for creating a stimulating environment for discussion. In particular, I thank Svetlana Stolpner for her openness to sharing her ideas. Special thanks also go to Parya MomayyezSiahkal, Emmanuel Piuze-Phaneuf, Maxime Boucher, and Herv\'{e} Lombaert for their help and prompt feedback. 

I am specially grateful to Diego Alejandro Macrini for allowing us to use his code to set up the matching framework for flux graphs. Furthermore, my heartfelt thanks go to my dear friend Babak Samari for helping me a lot with his insightful comments, and support.  


I would like to thank all of my friends in Montr\'{e}al city who have been supporting me on a daily basis: Mohammad Bagher Motevali, Danesh Ghafoorzadeh, Vahid Raissi Dehkordi, Danial Alizadeh, Bahareh Ghotbi, Sara Moghtadernejad, Abolfazl Mohebbi, and Sadegh Bolouki. You have made me feel at home!

Last, but not least, my deepest gratitude goes to my family for their unflagging love and support throughout my life; this dissertation would not be possible without them.  Thank you Maman, Baba, Mohammad, Mojtaba, Fereshteh, Azar, and Daee Arash. 

 % Secondly, I would like to express my greatest gratitude to my dear friend and colleague Babak Samari. Besides, I am so thankful to all my friends in Shape Analysis group who helped me a lot during my masters Parya Momayyez Siahkal, Svetlana Stolpner, Maxim Boucher, and Emmanuel Piuze. Thirdly, I would like to extend my thanks to my fellow friends at Center for Intelligent Machines who gave their best helps and supports to me:  Danial Alizadeh,  Vahid Raissi Dehkordi,  Afshin Taghvaeipoor, Bahareh Ghotbi, Nona Nobari, Majid Sheikholeslami, Kamran Ghaffari, Mehrsan Javan and all great friends I may miss their names here. Besides, I would like thank two of my best friends friends who made me glad and helped me with toughnesses during all these times: Sadegh Bolouki and Babak Samari. Finally, I am always thankful to my great family I have got who are all I have in this world. 