\documentclass[12pt,Bold,letterpaper,TexShade]{mcgilletdclass}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[dvips]{geometry}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algorithmic}
\addtolength{\hoffset}{0pt}                        
\addtolength{\voffset}{0pt}                        
\usepackage{amsthm}
\usepackage{rotating}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{indentfirst}
\usepackage{amsfonts}
\usepackage{amsmath}







\include{Title/title_page}

\makeindex[keylist]
\makeindex[abbr]

\listfiles%
\begin{document}

\maketitle%

\begin{romanPagenumber}{2}%

\SetDedicationName{\MakeUppercase{Dedication}}%
\SetDedicationText{\input{Dedication/dedication_page}}%
\Dedication%

\SetAcknowledgeName{\MakeUppercase{Acknowledgements}}%
\SetAcknowledgeText{\input{Acknowledgement/acknowledgment_page}}%
\Acknowledge%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         English Abstract                        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetAbstractEnName{\MakeUppercase{Abstract}}%
\SetAbstractEnText{ This thesis uses a frame fitting method developed in \cite{piuzephd} to describe the fiber geometry of mammalian hearts. Extending previous work, we analyze on a quantitative and qualitative level the extra information that this geometrical analysis can provide on infarcted hearts. Those hearts, right after the infarct, see the damaged region get filled with collagen and significantly lose contractile efficiency due to the loss of many fibers that were involved in the contraction during the beat cycle. General knowledge from Apparent Diffusion Coefficient (ADC) and Fractional Anisotropy (FA) map analysis does not provide an indication of how some surviving fibers can persist in the infarct region and how the remodeling process can allow some contractile function, although hindered and modified by the geometrical reorganization and the decreased connectivity between regions remote from the infarct and potential surviving fibers in the infarct region. Our connection form parameter calculation and frame fitting error plots allow us to visualize regions mostly filled with collagen within the infarct region, as well as regions where some fiber structure still shows up, demonstrating post infarct contractile potential after infarct, and how this new organization appears to be. Our approach could be a promising way to understand the process of remodeling of the heart wall following an infarct. This could also highlight the new geometrical structure and its relation to the unfortunately frequent secondary effects: tachycardia, arrhythmia and related abnormalities.}
\AbstractEn%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         French Abstract                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SetAbstractFrName{\MakeUppercase{ABR\'{E}G\'{E}}}%
\SetAbstractFrText{Ce m\'{e}moire se base sur une m\'{e}thode d'ajustement d\'{e}velopp\'{e}e dans \cite{piuzephd} pour d\'{e}crire la g\'{e}om\'{e}trie des fibres du coeur de mammif\`{e}res. En approfondissant le travail auparavant effectu\'{e}, nous analisons au niveau qualitatif et quantitatif les informations suppl\'{e}mentaires que nous apportent cette analise g\'{e}om\'{e}trique r\'{e}alis\'{e}e sur des coeurs avec infarctus. Ces coeurs, juste apr\`{e}s avoir subi un infarct, ont la r\'{e}gion touch\'{e}e par l'infarct qui se remplit de collag\`{e}ne et devient moins contractile de mani\`{e}re significative. En effet beaucoup des fibres impliqu\'{e}es dans la contraction pendant le cycle de battement du coeur sont mortes lors de l'infarct. Une connaissance g\'{e}n\'{e}rale d\'{e}livr\'{e}e par l'analise des cartographies d'anisotropie fractionnelle (FA) et de coefficient de diffusion apparent (ADC) ne donne aucune information sur d'\'{e}ventuelles fibres qui auraient survi \`{a} l'infarct et dans quelle mesure elles peuvent toujours jouer un r\^{o}le dans la fonction contractile m\^{e}me si celle-ci est gravement aggrav\'{e}e et modifi\'{e}e par la r\'{e}organisation g\'{e}om\'{e}trique des fibres et la moindre connectivit\'{e}e entre les r\'{e}gions \'{e}loign\'{e}es de l'infarct et les quelques r\'{e}gions dans la r\'{e}gion de l'infarct qui y ont surv\'{e}cu. Notre calcul de coefficients de connexions affines de Cartan et nos repr\'{e}sentations des erreurs d'ajustement de r\'{e}f\'{e}rentiel local nous permet d'avoir une visualisation int\'{e}r\'{e}ssante et informative des r\'{e}gions o\`{u} il persiste une structure des fibres qui d\'{e}montre le potentiel de contraction d'un coeur apr\`{e}s infarct, et comment cette r\'{e}organisation semble s'\^{e}tre effectu\'{e}e. Notre approche pourrait \^{e}tre un moyen prometteur de comprendre le processus de r\'{e}orgnanisation des fibres du coeur apr\`{e}s un infarctus. Elle pourrait \'{e}galement r\'{e}v\'{e}ler une nouvelle structure g\'{e}om\'{e}trique et sa relation \`{a} de malheureusement fr\'{e}quents effets secondaires: tachycardie, arrhythmie et autres anormalit\'{e}s li\'{e}es.}%
\AbstractFr%




\TOCHeading{\MakeUppercase{Table of Contents}}%
\LOTHeading{\MakeUppercase{List of Tables}}%
\LOFHeading{\MakeUppercase{List of Figures}}%
\tableofcontents %
\listoftables %
\listoffigures %

\end{romanPagenumber}

%\mainmatter %
 
\input{Chapters/Chapter1}
\input{Chapters/Chapter2}
\input{Chapters/Chapter3}
\input{Chapters/Chapter4}
\input{Chapters/Chapter5}


\bibHeading{References}
\bibliographystyle{plain}
\bibliography{myrefs}


\end{document}







